---
title: "New York Lottery Investigation"
author: "Jacob Mason"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will begin by pulling in data from my github repository.  This data was sorced from Data.gov and is assumed to be accurate.   There is data for 5 different New York Lotteries spanning different periods of time.  You must add raw to the begining of the url to pull just the csv file. 

```{r}
url1 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Cash_4_Life_Winning_Numbers__Beginning_2014.csv"

url2 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Mega_Millions_Winning_Numbers__Beginning_2002.csv"

url3 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Pick_10_Winning_Numbers__Beginning_1987.csv"

url4 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Powerball_Winning_Numbers__Beginning_2010.csv"

url5 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Take_5_Winning_Numbers__Beginning_1992.csv"


Cash4Life <- read.csv(url1)
MegaMillions <- read.csv(url2)
Pick10 <- read.csv(url3)
Powerball <- read.csv(url4)
Take5 <- read.csv(url5)
```

I have no idea the date ranges for these sets so lets take a look.  We will have to change the Draw.Date column in each data set to a date value and look at the greatest and lowest value. 

```{r}

summary(Cash4Life)
summary(MegaMillions)
summary(Pick10)
summary(Powerball)
summary(Take5)
```
This is a function to convert Draw Date into the proper date format and choose the earliest and latest date.
```{r}
summarize_draw_dates <- function(df, name = "Dataset") {
  # Check if 'Draw.Date' column exists
  if (!"Draw.Date" %in% names(df)) {
    cat("\n[", name, "] Missing 'Draw.Date' column.\n")
    return(NULL)
  }

  # Try multiple common formats just in case
  possible_formats <- c("%m/%d/%Y", "%Y-%m-%d", "%d-%m-%Y")
  converted <- NULL
  
  for (fmt in possible_formats) {
    try({
      converted <- as.Date(df$Draw.Date, format = fmt)
      if (all(!is.na(converted))) break
    }, silent = TRUE)
  }

  if (all(is.na(converted))) {
    cat("\n[", name, "] Could not parse dates. Please check the format.\n")
    return(NULL)
  }

  # Format dates as "mm/dd/yyyy"
  earliest <- format(min(converted, na.rm = TRUE), "%m/%d/%Y")
  latest <- format(max(converted, na.rm = TRUE), "%m/%d/%Y")

  cat("\n======", name, "======\n")
  cat("Earliest Draw Date:", earliest, "\n")
  cat("Latest Draw Date:", latest, "\n")
}


summarize_draw_dates(Cash4Life,name="Cash 4 Life")
```


