---
title: "New York Lottery Investigation"
author: "Jacob Mason"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


We will begin by pulling in data from my github repository.  This data was sorced from Data.gov and is assumed to be accurate.   There is data for 5 different New York Lotteries spanning different periods of time.  You must add raw to the begining of the url to pull just the csv file. 

```{r}
url1 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Cash_4_Life_Winning_Numbers__Beginning_2014.csv"

url2 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Mega_Millions_Winning_Numbers__Beginning_2002.csv"

url3 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Pick_10_Winning_Numbers__Beginning_1987.csv"

url4 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Powerball_Winning_Numbers__Beginning_2010.csv"

url5 <- "https://raw.githubusercontent.com/jacob-s-mason/New-York-Lottery-Numbers/main/Lottery_Take_5_Winning_Numbers__Beginning_1992.csv"


Cash4Life <- read.csv(url1)
MegaMillions <- read.csv(url2)
Pick10 <- read.csv(url3)
Powerball <- read.csv(url4)
Take5 <- read.csv(url5)
```

I have no idea the date ranges for these sets so lets take a look.  We will have to change the Draw.Date column in each data set to a date value and look at the greatest and lowest value. 

```{r}

summary(Cash4Life)
summary(MegaMillions)
summary(Pick10)
summary(Powerball)
summary(Take5)
```
This is a function to convert Draw Date into the proper date format and choose the earliest and latest date.
```{r}
summarize_draw_dates <- function(df, name = "Dataset") {
  # Check if 'Draw.Date' column exists
  if (!"Draw.Date" %in% names(df)) {
    cat("\n[", name, "] Missing 'Draw.Date' column.\n")
    return(NULL)
  }

  # Try multiple common formats just in case
  possible_formats <- c("%m/%d/%Y", "%Y-%m-%d", "%d-%m-%Y")
  converted <- NULL
  
  for (fmt in possible_formats) {
    try({
      converted <- as.Date(df$Draw.Date, format = fmt)
      if (all(!is.na(converted))) break
    }, silent = TRUE)
  }

  if (all(is.na(converted))) {
    cat("\n[", name, "] Could not parse dates. Please check the format.\n")
    return(NULL)
  }

  # Format dates as "mm/dd/yyyy"
  earliest <- format(min(converted, na.rm = TRUE), "%m/%d/%Y")
  latest <- format(max(converted, na.rm = TRUE), "%m/%d/%Y")

  cat("\n======", name, "======\n")
  cat("Earliest Draw Date:", earliest, "\n")
  cat("Latest Draw Date:", latest, "\n")
}
```


```{r}
summarize_draw_dates(Cash4Life,name="Cash 4 Life")
```



```{r}
print(head(Cash4Life))
```
Well that is not good.   I will need to split up the winning numbers column for more than likely all the data sets and the columns marked Winning Numbers.   Let me check the other to be sure.

```{r}
print(head(MegaMillions))

print(head(Pick10))

print(head(Powerball))

print(head(Take5))      
      
```
ALright, now to split up Winning Numbers
```{r}

MegaMillions_splt <- MegaMillions %>% separate(Winning.Numbers, into = paste0("Number", 1:5), sep = " ") %>% mutate(across(starts_with("Number"), as.integer))

print(head(MegaMillions))

MegaMillions_splt_long <- MegaMillions_splt %>%
  pivot_longer(
    cols = -Draw.Date,
    names_to = "Variable",
    values_to = "Value"
  ) %>% mutate(First_Integer = as.integer(str_extract(Value, "\\d"))) 

print(head(MegaMillions_splt_long))

MegaMillions_splt_long %>% 
  count(First_Integer) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  filter(!is.na(First_Integer)) %>%
  ggplot(aes(x = factor(First_Integer), y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste(n, "(", round(Percentage, 1), "%)", sep = "")), 
            vjust = -0.3, size = 3.5) +  # Label with count and percentage
  scale_x_discrete(limits = as.character(1:9)) +
  labs(
    title = "Distribution of First Digits",
    x = "Leading Digit",
    y = "Frequency"
  ) +
  theme_minimal()

```

THat was a wild goose chase.  Lets see a distribution of just the numbers.

```{r}



MegaMillions_splt_long %>%  filter(grepl("^Number", Variable)) %>%
  count(Value) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  filter(!is.na(Value)) %>%
  ggplot(aes(x = factor(Value), y = n)) +
  geom_col(fill = "steelblue") +  # Label with count and percentage
  scale_x_discrete() +
  labs(
    title = "Distribution of Values of MegaMillions Normal Balls",
    x = "Leading Digit",
    y = "Frequency"
  ) +
  theme_minimal()



```

Now something looks a little fishy looking at this distrobution...... Numbers after 54 were not chosen as frequently....  could be that these numbers were added at a later date in the data set.    Could be someting is off. 
 
```{r}


MegaMillions_splt_long %>%  filter(grepl("^Number5", Variable)) %>%
  count(Value) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  filter(!is.na(Value)) %>%
  ggplot(aes(x = factor(Value), y = n)) +
  geom_col(fill = "steelblue") +  # Label with count and percentage
  scale_x_discrete() +
  labs(
    title = "Distribution of Values of MegaMillions Normal Balls",
    x = "Leading Digit",
    y = "Frequency"
  ) +
  theme_minimal()


```

```{r}
summarize_draw_dates(MegaMillions_splt,name="Mega Millions")

MM_over52 <- MegaMillions_splt_long %>% filter(Value >50)

summarize_draw_dates(MM_over52 , "Over 52")
```
```{r}

```

